Progress report: \textit{Emys marmorata} sub-species identification and classification

### 2013-04-04 documents/turtle_report.Rnw
confusion matrices for multi. to do list.

\begin{itemize}
  \item multinomial logistic regression
  \item feed-forward neural networks
  \item random forests
\end{itemize}

@@

Dissimilarity based EAC estimated approximately \Sexpr{max(tmorph.dac)} optimal partitions. This corresponds to the number of partitions in both the Spinks et al. and shXXX classification schemes. However, the number of individuals assigned to each class is very different.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width = \textwidth]{figure/gm}
  \caption{<+caption text+>}
  \label{fig:gm}
\end{figure}

@@

I'm currently holding back on showing the tables because there are a lot of them.

But I can show the comparison of the predictive accuracies.

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/multi-map3}
    \label{fig:multi-map3}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/multi-map4}
    \label{fig:multi-map4}
  \end{subfigure}
  \caption{<+caption text+>}
  \label{fig:multi-map}
\end{figure}

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/nnet-map3}
    \label{fig:nnet-map3}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/nnet-map4}
    \label{fig:nnet-map4}
  \end{subfigure}
  \caption{<+caption text+>}
  \label{fig:nnet-map}
\end{figure}

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/rf-map3}
    \label{fig:rf-map3}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/rf-map4}
    \label{fig:rf-map4}
  \end{subfigure}
  \caption{<+caption text+>}
  \label{fig:rf-map}
\end{figure}

\section{Miscellaneous affairs}
\subsection{Evolution 2013}
Machine learning approaches for recognizing cyptic diversity: a case study using \textit{Emys marmorata}.

>>>>

\begin{itemize}
  \item multinomial logistic regression (Fig. \ref{fig:multi-map})
  \item feed-forward neural networks (Fig. \ref{fig:nnet-map})
  \item random forests (Fig. \ref{fig:rf-map})
\end{itemize}

@@

Dissimilarity based EAC estimated approximately \Sexpr{max(tmorph.dac)} optimal partitions (Table \ref{tab:dac}).

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width = \textwidth]{figure/gm}
  \caption{Visualization of PCA of \textit{E. marmorata}. The lower triangle is the pairwise comparison of the first three pricipal components. The upper left corner is the comparison of landmark dispersion for all specimens compared to the mean shape in red.}
  \label{fig:gm}
\end{figure}

@@

I'm currently holding back on showing the tables because there are a lot of them. I'll show the confusion matrix for the multinomial logicistic regressions because they are easy (Tables \ref{tab:multi-conf-sh1}, \ref{tab:multi-conf-sh2}, \ref{tab:multi-conf-sh3}, \ref{tab:multi-conf-spinks})

<<multi-tab, results = 'asis'>>=
mt <- lapply(tm.conf, function(x) xtable(as.table(x)))

for(ii in seq(length(mt))) {
  caption(mt[[ii]]) <- paste0('Confusion matrix for ',
                              names(mt)[ii],
                              ' classification scheme. Rows correspond to predicted class. Columns correspond to reference class.')
  label(mt[[ii]]) <- paste0('tab:multi-conf-', names(mt)[ii])
}

for(jj in seq(length(mt))) {
  print.xtable(mt[[jj]]
               , table.placement = 'ht')
}

@

But I can show the comparison of the predictive accuracies (Fig. \ref{fig:resamp}).

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/multi-map3}
    \label{fig:multi-map3}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/multi-map4}
    \label{fig:multi-map4}
  \end{subfigure}
  \caption{Geographic position of all turtles sampled. Both training and testing observations are plotted. Training set observations are circles while testing observations are larger squares. Testing set observations are classified based on a multinomial logistic regression model. \ref{fig:multi-map1}: sh1 classification scheme. \ref{fig:multi-map2}: sh2 classification scheme. \ref{fig:multi-map3}: sh3 classification scheme. \ref{fig:multi-map4}: spinks classification scheme.}
  \label{fig:multi-map}
\end{figure}

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/nnet-map3}
    \label{fig:nnet-map3}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/nnet-map4}
    \label{fig:nnet-map4}
  \end{subfigure}
  \caption{Geographic position of all turtles sampled. Both training and testing observations are plotted. Training set observations are circles while testing observations are larger squares. Testing set observations are classified based on a feed-forward single layer neural network model. \ref{fig:nnet-map1}: sh1 classification scheme. \ref{fig:nnet-map2}: sh2 classification scheme. \ref{fig:nnet-map3}: sh3 classification scheme. \ref{fig:nnet-map4}: spinks classification scheme.}
  \label{fig:nnet-map}
\end{figure}

@@

  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/rf-map3}
    \label{fig:rf-map3}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\textwidth}
    \centering
    \caption{}
    \includegraphics[width = \textwidth]{figure/rf-map4}
    \label{fig:rf-map4}
  \end{subfigure}
  \caption{Geographic position of all turtles sampled. Both training and testing observations are plotted. Training set observations are circles while testing observations are larger squares. Testing set observations are classified based on a random forest model. \ref{fig:rf-map1}: sh1 classification scheme. \ref{fig:rf-map2}: sh2 classification scheme. \ref{fig:rf-map3}: sh3 classification scheme. \ref{fig:rf-map4}: spinks classification scheme.}
  \label{fig:rf-map}
\end{figure}

\section{Future}
\begin{enumerate}
  \item remove juvelines
  \item look into multinomial logistic mixed-effects models
  \item other unsupervised methods, though that might have to wait for a follow up paper
    \begin{itemize}
      \item bayesian nonparametrics for categorical data
    \end{itemize}
\end{enumerate}

\section{Miscellaneous affairs}
\subsection{Evolution 2013}
How cryptic is cryptic diversity? Machine learning approaches to fine scale variation in morphology of \textit{Emys marmorata}.

### 2013-04-04 documents/turtle_report.tex
confusion matrices for multi. to do list.

\begin{itemize}
  \item multinomial logistic regression
  \item feed-forward neural networks
  \item random forests
\end{itemize}

\section{Preliminary results}
\subsection{Unsupervised}
Comparison of gap statistic over a very wide range of plausible partitions indicates that as the number of partitions increase, there is a marginal increasing gap statistic until approximately 31 after which there is a marginal decrease in gap statistic (Fig. \ref{fig:gap}). It is notable that the standard errors around the gap statistic values are very large, and the marginal increases in gap statistic with an increased number of partitions may not be important. Additionally, all gap statiic values are within 0.0065 of each other meaning that there is little over all consensus for how many clusters are present when comparing gap statistics.

@@

Dissimilarity based EAC estimated approximately 4 optimal partitions. This corresponds to the number of partitions in both the Spinks et al. and shXXX classification schemes. However, the number of individuals assigned to each class is very different.

% latex table generated in R 2.15.3 by xtable 1.7-0 package
% Wed Apr  3 19:43:09 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rr}
  \hline
 & tmorph.dac \\ 
  \hline
1 &  33 \\ 
  2 &   3 \\ 
  3 & 730 \\ 
  4 &  10 \\ 
   \hline
\end{tabular}
\caption{Number of specimens assigned to optimal number of partitions as determined by dissimilarity based EAC. Each column corresponds to a different partition, with the number assigned directly below it.}
\label{tab:dac}
\end{center}
\end{table}

\begin{figure}[ht]
  \centering
  \includegraphics[width = \textwidth]{figure/gm}
  \caption{<+caption text+>}
  \label{fig:gm}
\end{figure}

@@

I'm currently holding back on showing the tables because there are a lot of them.

But I can show the comparison of the predictive accuracies.

>>>>

\begin{itemize}
  \item multinomial logistic regression (Fig. \ref{fig:multi-map})
  \item feed-forward neural networks (Fig. \ref{fig:nnet-map})
  \item random forests (Fig. \ref{fig:rf-map})
\end{itemize}

\section{Preliminary results}
\subsection{Unsupervised}
Comparison of gap statistic over a very wide range of plausible partitions indicates that as the number of partitions increase, there is a marginal increasing gap statistic until approximately 40 after which there is a marginal decrease in gap statistic (Fig. \ref{fig:gap}). It is notable that the standard errors around the gap statistic values are very large, and the marginal increases in gap statistic with an increased number of partitions may not be important. Additionally, all gap statiic values are within 0.0064 of each other meaning that there is little over all consensus for how many clusters are present when comparing gap statistics.

@@

Dissimilarity based EAC estimated approximately 5 optimal partitions (Table \ref{tab:dac}).

% latex table generated in R 2.15.3 by xtable 1.7-0 package
% Thu Apr  4 14:25:03 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rr}
  \hline
 & tmorph.dac \\ 
  \hline
1 & 722 \\ 
  2 &   1 \\ 
  3 &   1 \\ 
  4 &   1 \\ 
  5 &   1 \\ 
   \hline
\end{tabular}
\caption{Number of specimens assigned to optimal number of partitions as determined by dissimilarity based EAC. Each column corresponds to a different partition, with the number assigned directly below it.}
\label{tab:dac}
\end{center}
\end{table}

\begin{figure}[ht]
  \centering
  \includegraphics[width = \textwidth]{figure/gm}
  \caption{Visualization of PCA of \textit{E. marmorata}. The lower triangle is the pairwise comparison of the first three pricipal components. The upper left corner is the comparison of landmark dispersion for all specimens compared to the mean shape in red.}
  \label{fig:gm}
\end{figure}

@@

I'm currently holding back on showing the tables because there are a lot of them. I'll show the confusion matrix for the multinomial logicistic regressions because they are easy (Tables \ref{tab:multi-conf-sh1}, \ref{tab:multi-conf-sh2}, \ref{tab:multi-conf-sh3}, \ref{tab:multi-conf-spinks})

% latex table generated in R 2.15.3 by xtable 1.7-0 package
% Thu Apr  4 14:25:03 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrr}
  \hline
 & CCR & marm & pall \\ 
  \hline
CCR &  40 &  14 &  11 \\ 
  marm &  25 &  74 &   2 \\ 
  pall &   3 &   1 &  10 \\ 
   \hline
\end{tabular}
\caption{Confusion matrix for sh1 classification scheme. Rows correspond to predicted class. Columns correspond to reference class.}
\label{tab:multi-conf-sh1}
\end{center}
\end{table}
% latex table generated in R 2.15.3 by xtable 1.7-0 package
% Thu Apr  4 14:25:03 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrr}
  \hline
 & CCR & marm & pall \\ 
  \hline
CCR &  44 &  12 &  10 \\ 
  marm &  17 &  82 &   6 \\ 
  pall &   2 &   1 &   7 \\ 
   \hline
\end{tabular}
\caption{Confusion matrix for sh2 classification scheme. Rows correspond to predicted class. Columns correspond to reference class.}
\label{tab:multi-conf-sh2}
\end{center}
\end{table}
% latex table generated in R 2.15.3 by xtable 1.7-0 package
% Thu Apr  4 14:25:03 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrrr}
  \hline
 & CCR & marm & pall & SJ \\ 
  \hline
CCR &  37 &  11 &   9 &   2 \\ 
  marm &  18 &  59 &   3 &  10 \\ 
  pall &   7 &   2 &  10 &   0 \\ 
  SJ &   1 &   4 &   1 &   7 \\ 
   \hline
\end{tabular}
\caption{Confusion matrix for sh3 classification scheme. Rows correspond to predicted class. Columns correspond to reference class.}
\label{tab:multi-conf-sh3}
\end{center}
\end{table}
% latex table generated in R 2.15.3 by xtable 1.7-0 package
% Thu Apr  4 14:25:03 2013
\begin{table}[ht]
\begin{center}
\begin{tabular}{rrrrr}
  \hline
 & 1 & 2 & 3 & 4 \\ 
  \hline
1 & 107 &   9 &  13 &  13 \\ 
  2 &   1 &   8 &   0 &   0 \\ 
  3 &   4 &   0 &   8 &   3 \\ 
  4 &   4 &   0 &   2 &   8 \\ 
   \hline
\end{tabular}
\caption{Confusion matrix for spinks classification scheme. Rows correspond to predicted class. Columns correspond to reference class.}
\label{tab:multi-conf-spinks}
\end{center}
\end{table}

But I can show the comparison of the predictive accuracies (Fig. \ref{fig:resamp}).
